---
title: "PLSC30500, Fall 2022"
subtitle: "Week 1. Programming, pt. 2"
# author: "Andy Eggers & Molly Offer-Westort"
output: 
  xaringan::moon_reader:
    self_contained: true
    css: [default, uchicago_pol_meth.css]
    nature:
      highlightLines: true

---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(fig.height=5, fig.width=8.5, fig.align="center", warning = F, message = F)
```


class: inverse, middle, center

# A bit of base `R`

---

# Assignment via `->`

```{r}
8
x <- 8
x
x^2
y <- x^2
y
```


---

## Vectors, functions, data frames

```{r}
x <- c(1, 2, 5)
mean(x)
max(x)
median(x)
x^2
data.frame(a = x, b = x^2)
```

---

# Adding new columns

```{r}
x <- c(1, 2, 5)
dat <- data.frame(a = x, b = x^2)
dat$c <- dat$a + 3
dat
```

---

# The `sample()` function

```{r}
x <- c(1, 2, 5)
sample(x) # reshuffle
sample(x, size = 1) # draw one
sample(x, size = 2, replace = F) # draw two without replacement
sample(x, size = 5, replace = T) # draw five with replacement
sample(x, 10, T, prob = c(.05, .2, .75)) # unequal probs
```

---

# Normal distribution

```{r}
rnorm(n = 5, mean = 0, sd = 1) # random draws
dnorm(x = 0, mean = 0, sd = 1) # density
dnorm(x = 4, 0, 1)
pnorm(q = 1.96, 0, 1) # cumulative density
pnorm(0, 0, 1)
```

---

# Make a sequence

```{r}
1:10
seq(from = 1, to = 10, by = 1)
seq(from = 1, to = 10, by = 2)
```


---

# The pipe: `|>`

```{r}
x
```

Equivalent code: 

```{r}
mean(x)
c(1,2,5) |>
  mean()
```

You will also see `%>%` instead of `|>`. 

`Ctrl-shift-m` in RStudio (can adjust preferences to determine which one appears).


---

class: inverse, middle, center

# `tidyverse` for data wrangling: an introduction 

---

# Tidyverse and base `R` 

For many tasks, `R` offers many possible solutions. 

`tidyverse` is a "package of packages" (including `ggplot`) with a more-or-less consistent programming style.

We will tend to use `tidyverse` approaches.

---

## `mutate()` for adding variables

```{r}
tibble(a = seq(0, 5, by = .25)) |>  # data.frame() also works
  mutate(b = a^2) |> 
  ggplot(mapping = aes(x = a, y = b)) + # note '+'
  geom_point()
```


---

## `filter()` for choosing rows

```{r}
tibble(a = runif(n = 10000)) |>
  filter(a <= .3) |> 
  ggplot(mapping = aes(x = a)) + 
  geom_histogram(breaks = seq(0, .3, by = .01))
```

---

## `select()` for choosing columns

```{r}
dat <- tibble(a = c(1, 2, 5)) |>
  mutate(b = a^2,
         c = dnorm(a),
         d = 1)

dat |> 
  select(a, c)

dat |> 
  select(-c)
```

---

## Grouped summaries

```{r}
dat <- tibble(a = c(1, 1, 2, 5)) |>
  mutate(b = runif(4))

dat

dat |> 
  group_by(a) |> 
  summarize(c = mean(b))
```


---

class: inverse, middle, center

# An application with real data

---


# Load V-Dem dataset

1. Install the installer: `install.packages("devtools")`
1. Install the package from github: `devtools::install_github("vdeminstitute/vdemdata")`
1. Load the package: `library(vdemdata)`

```{r echo = F}
library(vdemdata)
```


Now by typing `vdem` you can access the dataset.


???

Can also get necessary subset on repository: data/vdem_long_extract.csv

Note one objective here is just to practice the skills on political science data rather than e.g. `mpg` or `diamonds`.


---

# Getting a feel for the data (1)

`vdem` is big! 

```{r}
dim(vdem)
```

--

Typing `vdem` is overwhelming and unhelpful; `View(vdem)` better but takes a while to load. 

---

#### Look at column names: 

.small-output[
```{r}
names(vdem)  # colnames(vdem) also works
```
]

---


#### `tibble`s: better viewing

<!-- see
https://github.com/yihui/xaringan/issues/169
for shortcut
--> 

.small-output[
```{r}
library(tidyverse)
vdem |>  
  as_tibble()   # "tibble" package is part of "tidyverse"
```
]



---

### Choose columns with `select()`


```{r}
vd <- as_tibble(vdem) |>   
  select(country_name, year, e_regionpol_6C,  #<<
         v2x_polyarchy, e_gdppc)  #<<
vd
```

---


### Choose rows with `filter()`

```{r}
vd |>  
  filter(year == 2015) #<<
```

---

### Create new variables with `mutate()`

```{r}
vd |>  
  filter(year == 2015) |>  
  mutate(polyarchy_times_100 = 100*v2x_polyarchy) |>  #<<
  select(country_name, v2x_polyarchy, polyarchy_times_100)
```

---

### Coding exercise: reverse-engineer the plot!

```{r rev-eng-0, echo = F, fig.height=5.5, fig.width=9}
vd |>  
  filter(country_name %in% c("Iraq", "France", "China", #<<
                             "United States of America")) |>   #<<
  ggplot(aes(x = year, y = v2x_polyarchy, col = country_name)) +
  geom_line()
```

???

Polyarchy is a multiplicative ("necessary conditions") index with five components: 

1. Elected officials
2. Clean elections
3. Associational autonomy
4. Inclusive citizenship
5. Freedom of expression and alternative sources of information

---

```{r ref.label = "rev-eng-0", fig.height=5.5, fig.width=9}
```

---

### Coding exercise: reverse-engineer the plot!

```{r rev-eng-1, echo = F, fig.height=5.5, fig.width=9, message = F}
vd |> 
  group_by(e_regionpol_6C, year) |>  
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) |>  
  ggplot(aes(x = year, y = mean_polyarchy, 
             col = factor(e_regionpol_6C))) +
  geom_line() + 
  scale_color_discrete(labels = c("E. Europe & Central Asia",
                                "Latin America & Caribbean",
                                "Middle East & N. Africa",
                                "Sub-Saharan Africa",
                                "W. Europe & N. America",
                                "Asia & Pacific")) +
  labs(color = "Region",
       caption = "Regions from e_regionpol_6C (V-Dem)",
       x = "Year",
       y = "Average polyarchy score in region")
```
---

## First task: compute average polyarchy score within region over time

First for the whole dataset:
```{r}
vd |>  
  group_by(e_regionpol_6C) |>  
  summarize(mean_polyarchy = mean(v2x_polyarchy))
```

--

Oops! 


---

## Try again: compute average polyarchy score within region over time

First for the whole dataset:
```{r}
vd |> 
  group_by(e_regionpol_6C) |> 
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) #<<
```

---

## Getting closer:

Now for a single year:

```{r}
vd  |>  
  filter(year == 2015) |>   #<<
  group_by(e_regionpol_6C) |>  
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) 
```

---

## Average polyarchy score within region over time 

```{r}
vd |> 
  group_by(e_regionpol_6C, year) |>   #<<
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) 
```

---

### Now to plot it

```{r, fig.height=4, fig.width=9}
vd |>  
  group_by(e_regionpol_6C, year) |> 
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) |>  
  ggplot(aes(x = year, y = mean_polyarchy, col = e_regionpol_6C)) + 
  geom_line()
```

Hmm.

---

We need to tell `ggplot` that `e_regionpol_6C` is a categorical variable, not a continuous variable.

```{r, fig.height=4, fig.width=9}
vd |>  
  group_by(e_regionpol_6C, year) |> 
  summarize(mean_polyarchy = mean(v2x_polyarchy, na.rm = T)) |>  
  ggplot(aes(x = year, y = mean_polyarchy, col = factor(e_regionpol_6C))) + #<< 
  geom_line()
```


---

#### The rest is fiddly stuff with labels. 


```{r ref.label = "rev-eng-1", fig.height=5.5, fig.width=9, eval = F}
```
---



### Alternative way to do something similar


```{r fig.width=9}
vd |>  
  filter(year > 1900) |>  
  ggplot(aes(x = year, y = v2x_polyarchy,
             col = factor(e_regionpol_6C))) +  #<<
  geom_smooth(span = .5)  #<<
```


---


### Plotting grouped summaries (2)


```{r fig.width=10.5}
vd |>  
  filter(year > 1900) |>  
  group_by(year) |> 
  summarize(mean_poly = mean(v2x_polyarchy, na.rm = T),
            mean_gdppc = mean(e_gdppc, na.rm = T)) |>  
  mutate(my_size = ifelse(year %in% c(1946, 1960, 1989, 2018), 8, 0)) |>  
  ggplot(aes(x = mean_gdppc, y = mean_poly)) +  #<<
  geom_path() + 
  geom_point() + 
  geom_text(aes(label = year, size = my_size), show.legend=F)
```


