---
title: "PLSC30500, Fall 2022"
subtitle: "Week 3. Summarizing distributions"
# author: "Andy Eggers & Molly Offer-Westort"
output: 
  xaringan::moon_reader:
    self_contained: true
    css: [default, uchicago_pol_meth.css]
    nature:
      highlightLines: true

---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(fig.height=5, fig.width=8.5, fig.align="center", warning = F, message = F)
```


# Expected value: discrete case 

For discrete R.V. with probability mass function (PMF) $f$, the *expected value* of $X$ is 

$$E[X] = \sum_x x f(x) $$

Could write 

$$E[X] = \sum_{x \in \text{Supp}[X]} x f(x) $$

---

# Example 

| $x$  | $P(X = x)$ |
|------|:------------:|
|   0   |     .2     |
|   1   |     .5     |
|   3   |     .3     |

What is $E[X] = \sum_x x f(x)$?

--

$$\begin{aligned}
E[X] &= \sum_x x f(x) \\
 &= 0 \times .2 + 1 \times .5 + 3 \times .3 \\
 &= 1.4
\end{aligned}$$


---

# Illustration

```{r, echo = F}
dat <- tibble(x = c(0, 0, 1, 1, 3, 3), y = c(0, .2, 0, .5, 0, .3), item = c(1, 1, 2, 2, 3, 3), type = rep(c("bottom", "top"), 3)) 

dat|> 
  ggplot(aes(x = x, y = y, group = item)) + 
  geom_line() + 
  geom_point(data = filter(dat, type == "top")) + 
  expand_limits(y = .65) +
  geom_vline(xintercept = .5*1 + .3*3, lty = 2)
```

---

# E[X] as X's center of mass 

Suppose we place a weight $f(x)$ at each value $x \in \text{Supp}(X)$ along a weightless rod.

--

Where is the **center of mass**, i.e. the point where the rod balances?

--

It is the point $c$ where $\sum_x (x - c) f(x) = 0$.

--

That point is $E[X]$.

--

**Proof**:

$$\begin{aligned} \sum_x (x - E[X]) f(x) &= \sum_x \left( x f(x) - E[X] f(x) \right) \\
&= \sum_x x f(x) -  \sum_x E[X] f(x) \\
 &= E[X] - E[X] \sum_x f(x) \\
 &= E[X] - E[X] \\
 &= 0
\end{aligned}$$

---

# E[X] as a "good guess"

Define **mean squared error** of a R.V. $X$ about $c$:

$$ E[(X - c)^2] $$

--

$E[X]$ is the choice of $c$ that minimizes MSE. (Theorem 2.1.24, illustrated on homework.)

---

# Expectation vs average

Given a set of numbers $x_1, x_2, \ldots, x_n$ we can take the **average**: 

$$ \overline{x} = \frac{1}{n} \sum_{i = 1}^n x_i $$

What is the connection between $E[X] = \sum_x x f(x)$ and $\overline{x} = \frac{1}{n} \sum_{i = 1}^n x_i$?

--

- $E[X]$ summarizes a random variable $X$; $\overline{x}$ summarizes a set of numbers
- if each $x_i$ is an (iid) **sample** from $X$, $\overline{x}$ approximates $E[X]$
- if each $x$ appears with frequency $f(x)$, then $\overline{x} = E[X]$

---

# Expectation vs average (2)

Given PMF:

| $x$  | $P(X = x)$ |
|------|:------------:|
|   0   |     .2     |
|   1   |     .5     |
|   3   |     .3     |

Then $E[X] = 0 \times .2 + 1 \times .5 + 3 \times .3 = 1.4$.

--

Can also compute as the mean of a vector in which each $x$ appears with frequency $f(x)$: 

```{r}
x <- c(0, 0, 1, 1, 1, 1, 1, 3, 3, 3)
mean(x)
```

--

May clarify why, in any given sample, $\overline{x} \neq E[X]$.

---

# The continuous case 

For continuous R.V. $X$, 

$$ E[X] = \int_{-\infty}^{\infty} x f(x) dx $$ 

--

```{r, echo = F}
xs <- seq(-.5, 2, by =.01)
ys <- dnorm(xs)

# get expectation numerically -- two options

# weighted sum of xs in my sequence
ex1 <- sum(xs*ys)/sum(ys)

# average of a sample
x_samp0 <- rnorm(100000)
x_samp <- x_samp0[x_samp0 > -.5 & x_samp0 < 2]
ex2 <- mean(x_samp)

# plot it
tibble(x = c(-.5, xs, 2), y = c(0, ys, 0)) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_polygon(fill = "darkgray") + 
  theme_bw() + 
  geom_vline(xintercept = ex1, col = "red") # +
  # geom_vline(xintercept = ex2, col = "blue") # 

```

???

Just convey idea that balance will still work, we won't be doing integration.

And we will stick to the discrete case as much as possible.

---

# Expectation of a function of an R.V.

Let $g(X)$ be a function of discrete random variable $X$. It is a random variable with PMF $f(g(X))$. 

(e.g. $g(X) = 3X$)


Then $$ E[g(X)] = \sum_{x \in \text{Supp}(X)} g(x) f(x)$$

<!-- $$\begin{aligned} -->
<!-- E[g(X)] &= \sum_{x \in \text{Supp}(g(X))} g(X) f(g(x)) \\ -->
<!-- &= \sum_{x \in \text{Supp}(X)} g(x) f(x) \\ -->
<!-- \end{aligned}$$. -->


--


Why is this the same as $\sum_{x \in \text{Supp}(g(X))} g(X) f(g(x))$?

| $x$  | $P(X = x)$ | $g(x)$ | $P(g(X) = g(x))$
|------|:------------:|------|:-------:|  
|   0   |     1/3     |  0 |  1/3 |
|   1   |     2/3    |  3 |  2/3 |





